---
title: "Activity Monitoring"
author: "Yash Singhal"
output:
  html_document:
    theme: cosmo
---

### Loading and preprocessing the data

```{r include=FALSE}
library(dplyr)
library(zoo)
library(ggplot2)
```


```{r}

df <- read.csv("activity.csv")

```


Here I am Adding a Column using the Time Series Function (Day elapsed Column)

```{r}
as.ts(df$date) -> df$Time_Series
```

### What is mean total number of steps taken per day?

Using Group by operation on date column to determine the total no of steps each date

```{r}
df %>% group_by(date) %>% summarise(Steps=sum(steps,na.rm=T)) -> L
paste("Total Steps taken per day",Total_Number <- sum(df$steps,na.rm = T)/max(df$Time_Series))
paste("Median no of steps taken per day",median(L$Steps))

```

Histogram of the total number of steps taken each day

```{r}
hist(L$Steps,main="Total no of steps Taken each day",col="red")
```

### What is the average daily activity pattern?

Adding a Column

```{r}

mean(L$Steps) -> Hm
df$Average <- df$steps/Hm
```

LineSeriesPlot

```{r}
with(na.omit(df),plot(x=interval,y=Average,type="l",col="red",main = "Average Steps 5 minute interval"))

```

That 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

```{r}
df[which.max(df$Average),]
```

### Imputing missing values

```{r}
paste("No of missing value",colSums(is.na(df))[1])
```

Median would be better suited to fill the missing values as dataset is heavily skewed

```{r}
hist(df$steps)
```

Filling in the median

```{r}
median(df$steps,na.rm=T) -> Median
df1 <- df
na.fill(df$steps,0) -> df1$steps
```

Grouping with respect to new data set

```{r}
df1 %>% group_by(date) %>% summarise(Steps=sum(steps)) -> L1
```

```{r}
paste("Total Steps taken per day",Total_Number <- sum(df1$steps,na.rm = T)/max(df$Time_Series))
paste("Median no of steps taken per day",median(L1$Steps))
hist(L1$Steps,col="red")
```

Almost No difference since the Median was 0

### Are there differences in activity patterns between weekdays and weekends

Making a Function to Add column whether there is a weekend or weekday.New Factor Column is DayT

```{r}
Check_Weekday <- function(x){
seq <- c("Monday","Tuesday","Wednesday","Thursday","Friday")
if (x %in% seq){return ("weekday")}
else {return ("weekend")}
}
sapply(weekdays(as.POSIXct(df1$date)),Check_Weekday) -> df1$DayT
as.factor(df1$DayT) -> df1$DayT

```

Adding a Column and Plotting Panels

```{r}
df1$steps/Hm -> df1$Average

ggplot(aes(x=interval,y=Average),data=df1)+geom_line(col="blue")+facet_wrap(~DayT,nrow = 2)
```

